### 2024-02-02 analysis/paper/paper.qmd
work on result

---
title: "Characterization of the microbial and fungal communities in ancient pottery from Neolithic southwestern China"
author:
  - Li-Ying Wang:
      correspondence: "yes"
      email: liying15@gmail.com
      orcid: 0000-0001-9060-2125
      institute:
        - AS
  - Kuei-Chen Lin:
      institute: AS
      orcid: 0000-0003-1689-0558
  - Zhiqing Zhou:
      orcid: 0000-0003-1689-0551
      institute:
        - name: Chengdu Institute of Cultural Relics and Archaeology
  - Yin-Ru Chiang:
      correspondence: "yes"
      institute: AS
      orcid: 0000-0003-1689-0559
institute:
  - AS:
      name: Academia sinica
      address: 130, Section 2, Academia Rd, Nangang District, Taipei City, Taiwan
title-block-published: "Last updated"  
date: now
date-format: long
format: 
  docx:
    reference-doc: "../templates/template.docx" # Insert path for the DOCX file
execute:
  echo: true
  warning: false
  message: false
  comment: "#>"
  fig-path: "../figures/"
  fig-dpi: 600
filters:
  - ../templates/scholarly-metadata.lua
  - ../templates/author-info-blocks.lua
  - ../templates/pagebreak.lua
bibliography: references.bib
csl: "../templates/journal-of-archaeological-science.csl" # Insert path for the bib-style
abstract: |
  Text of abstract
keywords: |
  keyword 1; keyword 2; keyword 3
highlights: |
  These are the highlights. 
---

<!-- This is the format for text comments that will be ignored during renderings. Do not put R code in these comments because it will not be ignored. -->

<!-- With the following code you can access and display values from the yml header above. -->

Keywords: `r rmarkdown::metadata$keywords`

Highlights: `r rmarkdown::metadata$highlights`

<!-- The actual document text starts here: -->

```{r setting}
#| echo: false
# Format integers
options(scipen=999)
knit_print.integer = function(x, ...) {
  prettyNum(x, big.mark=",")
}

registerS3method(
  "knit_print", "integer", knit_print.integer,
  envir = asNamespace("knitr")
)

```

# Introduction

Ceramics are the most common artifacts unearthed from archaeological sites, offering valuable insights into diverse anthropological facets encompassing cultural, social, and economic dimensions. Foodway is one of the topics. Food residues absorbed into the matrices of ceramics are useful sources to explore pottery use and past culinary practices. This analytical pursuit heavily relies on contemporary scientific methodologies. For example, organic residues, particularly lipids, extracted from pottery through a series of chemical techniques enable the identification of terrestrial animals, marine resources, and plants [@Evershed2008; @Evershed2002; @Evershed2008marine; @Craig2011; @Heron2016]. Furthermore, the identification of foodstuffs contained in storage vessels can be achieved through ancient DNA analysis, providing direct evidence of pottery content at the genes or species levels [@Foley2012; @Hansson2008; @Robinson2017]. These analytical approaches underscore the significance of preserving the condition of food residues within archaeological contexts, ensuring the efficacy of these scientific methodologies.

Recently, the application of next-generation sequencing in microbiology has brought a novel biomolecular approach to examine ancient human activity through microorganisms present in archaeological samples, including human remains [@Cano2000; @Ubaldi1998; @Ckimsis2023], teeth [@Adler2013; @Warinner2015; @De2013; @Preus2011], archaeological deposits, and minerals [@Margesin2017; @Siles2018; @Lenehan2017; @Boivin2023]. Such microbial archaeology could also be useful for studying food residues adhering to or absorbed into pottery matrix. Based on the assumption that food spoilage mainly results from microbial activity, the microorganisms preserved in pottery within burial contexts could be related to the last prehistoric use of the pot, such as food that was cooked in it. For example, microorganisms including *Proteus spp.*, *Weissella spp.*, *Pseudomonas spp.*, and *Enterobacteriaceae* are frequently associated with meat [@Tassew2010; @Stellato2016], while *Aspergillus spp.*, *Penicillium spp.*, *Fusarium spp.*, *Rhizopus spp.* are commonly found in association with rice [@Li2022; @Bertuzzi2019; @Tahir2012]. Moreover, *Acetobacter*, *Lactobacillus*, *Bacillus*, *Bifidobacterium* are indicative of fermented foods [@Tamang2016; @Xiao2020; @Zhang2018], and *Paenibacillus*, *Chryseobacterium* are largely found in mixed rice-vegetable meals [@Helmond2017; @Nde2021]. Given the potential widespread occurrence of microbial communities across various food categories, more studies are still needed to understand the characteristics of microbial communities in ancient pottery.

In this study, our objective is to investigate the feasibility of using microbial communities as indicators of past food processing in pottery by focusing on the potsherds from the Guijiabao site in southwest China. As a long-term occupation site from Neolithic period to historical period, Guijiabao revealed abundant archaeological assemblages and remaines to explore the adaptation of farming communities in a mountain setting. The post-excavation analysis of Guijiabao is still in progress. This study serves as a pilot study to examine the pottery use from a perspective of microbial archaeology. We aim to explore the differences in overall microbial communities, including bacterial and fungal communities, through a comparative analysis of different parts of ceramic samples and soil samples. Our main research question is: Can microbial compositions serve as evidence for distinguishing various food categories based on the prevalence of spoilage microorganisms? We hypothesized that the microbial composition in absorbed residues within the pottery matrix will exhibit distinctions relevant to food residues when compared to those on the surface of pottery and in surrounding soils. While microorganisms offer indirect evidence of past food practices, our focus is on their potential to provide valuable information that complements existing scientific approaches to analyzing food residues. This innovative perspective could shed light on food practices from a microbial perspective that has not been fully discussed in archaeology.

# Site Description and Sampling

Guijiabao is located in the Yanyuan Basin in the middle Hengduan Mountains of southwest China. The chronological framework of Guijiabao includes the Neolithic period (5000-3700 BP), the Bronze Age (3200-2700 BP), and the historical period spanning from the late North Song to early Ming dynasties (1200-500 BP) based on radiocarbon dates [@Hao2022; @Huan2022]. Given its location along the route of population movements, Guijiabao has been a place for cultural exchanges with neighboring regions, including the northern and southern parts of the Hengduan Mountains, the Sichuan Basin, and western Yunnan since the Neolithic period. This culture interactions is evident in the similarities observed in pottery styles and decorations [@Guijiabao2017a]. Excavation at Guijiabao revealed various pottery forms, including pots, jars, urns, bowls, and basins. Predominantly, the pottery types are characterized by grayish high-tempered fine ware and reddish-brown coarse ware featuring cord-marked or basket-impressed patterns [@Hao2022; @Guijiabao2017a].

As the earliest Neolithic site in the Yanyuan Basin, Guijiabao provides valuable data for exploring the lives of people who have lived in a highland environment (approximately 2400 m above sea level). Archaeobotanical evidence and phytolith analyses suggested that mixed crop agriculture, specifically rice and millet, was introduced to this region since the occupation of the site [@Huan2022; @Guijiabao2016]. Cultivated crops, such as broomcorn millet, foxtail millet, and rice were identified as the primary agricultural practices at the site [@Deng2019; @Guijiabao2016]. Hunting and fishing could be also part of local subsistence activity, according to the finding of stone hunting implements and net sinkers [@Hao2022; @Guijiabao2020a]. Examining pottery used in daily life becomes critical in shedding light on food practices, offering more evidence for understanding local subsistence via the lens of food preparation.

As a preliminary investigation of the feasibility of using a microbial approach for detecting evidence of cooking and food types, we sampled five potsherds of typical ceramics from Guijiabao, with a pottery sample from Gaoshan, anorther Neolithic site in southwest China, as a comparison. Gaoshan is located in southwest Chengdu Plain, an alluvial plain in Sichuan, and dates back to the early Baodun period, the earliest Neolithic culture (ca. 4700-3700 BP) in the Chengdu Plain [@Gaoshan2017]. Archaeological evidence from the Gaoshan site suggests a primary subsistence pattern involving rice cultivation and pig husbandry, supplemented by millet cultivation [@Gaoshan2017; @JiangGaoshan2017; @Lee2020; @He2020]. In addition to the pottery samples, we collected three soil samples from cultural layers across various excavation units at Guijiabao, alongside two soil samples from Gaoshan.
After sampling, the potsherds were securely stored in sealed boxes, while soil samples were preserved in sterile vials, both of which were subsequently frozen at -80 °C in preparation for DNA analyses.

# Methods

To minimize potential sources of contamination from the environments, we cleaned all lab benches, equipment, and pipettors with 3% bleach, followed by 99% ethanol and UV light. The sample preparation and DNA extraction were conducted in the biomolecular laboratory facilities at Academia Sinica, Taiwan. Because modern contamination could be a significant concern when interpreting the data, we used negative controls during sample preparation, and reagent blanks for DNA extraction, and PCR amplification to monitor extraneous DNA from surroundings, reagents, and consumables. A total of 12 powdered pottery samples (interior and exterior parts), 5 soil samples from Guijiabao and Gaoshan, and 3 negative control samples were analyzed in this study.

## Sample preparation and DNA extracation

Each pottery sample was divided into two portions: the interior, representing the middle part of the potsherd matrix, and the exterior surface exposed to the environment. This division aimed to facilitate a comparative analysis of the microbial communities in residues absorbed into the potsherds versus those on the surface. Approximately 1.5 g of each sample was collected using a sterile drill operated at a low speed to prevent heat generation. The powdered sample was initially gathered on pre-combustion foil and immediately transferred to a sterile Eppendorf tube. All the tools were cleaned between each iteration to minimize cross-contamination. For the soil samples, around 1g was collected using sterile spatulas and transferred to a sterile Eppendorf tube.

Total DNA from each sample was extracted in a portion of 300 mg using the DNeasy PowerSoil Pro kit (Qiagen, Germany), and the DNA extracts from the same sample were pooled. The concentration of DNA was measured using a Qubit 2.0 fluorometer and standardized to 1 ng μL^−1^.

## PCR analyses

Given the fragmentation of DNA in archaeological contexts, we targeted single hypervariable regions, V1, V4, and V6 of 16S rRNA, and ITS1 of 18S rRNA, to cover both bacterial and fungal communities. The V1 region was amplified using the primers 8F (5'-AGAGTTTGATYMTGGCTCAG-3') and 120R (5'-TTACTCACCCGTICGCCRCT-3') [@Adler2013], the V4 region was amplified using arch519F (5′-CAGCMGCCGCGGTAA-3′) and arch806R (5′-GGACTACVSGGGTATCTAAT-3′), and the V6 region was amplified using 926F (5’-AAACTYAAAKGAATTGACGGG-3’) and 1046R (5’-CGACARCCATGCASCACCT-3’) [@Adler2013; @Siles2018]. For fungal communities, the region ITS1 of 18S rRNA was amplified using the primers ITS1F (5’-CTTGGTCATTTAGAGGAAGTAA-3’) and ITS2 (5′-GCTGCGTTCTTCATCGATGC-3′) [@Siles2018]. Barcoded forward adapters were added to the PCR primers to facilitate sample pooling for library preparations. The DNA amplification for bacteria was conducted using the Taq DNA polymerase Master Mix RED (Ampliqon) under the following program: initial denaturation at 95 °C for 1 min, followed by 35 cycles of denaturation at 95 °C for 30 s, primer annealing at 50 °C for 30 s, extension at 72 °C for 30 s, and final elongation at 72 °C for 5 min. For fungi, the thermal conditions are the same as for bacteria, with the exception of 40 cycles and primer annealing temperature at 57 °C for 40 s.

The DNA fragments in PCR products were purified with the GenepHlow gel extraction kit  (Geneaid Biotech Ltd.) after conducting 2 % agarose gel electrophoresis. The amplified products were then pooled and used to build a DNA library using the Illumina TruSeq DNA library preparation protocol.

## DNA library construction and sequencting

## Taxonomic classification

# Results

```{r get-data}
#| label: get-data
#| echo: false
#| cache: true
library(tidyverse)
library(readxl)
# read original count data
OTU_V4 <- read.table(here::here("analysis","data","raw_data","V4_asv_table.txt"), header = T, sep="\t")
OTU_V1 <- read.table(here::here("analysis","data","raw_data","V1_asv_table.txt"), header = T, sep="\t")
OTU_V6 <- read.table(here::here("analysis","data","raw_data","V6_asv_table.txt"), header = T, sep="\t")
OTU_ITS1 <- read.table(here::here("analysis","data","raw_data","ITS1_asv_table.txt"), header = T, sep="\t")

OTU_reads_total <- function(df) {
  sum_read <- df %>%
    rowwise() %>% 
    mutate(sum_read = sum(across(where(is.numeric)))) %>% 
    select(sum_read)
  return(sum(sum_read))
}

V4_total_reads <- OTU_reads_total(OTU_V4)
V1_total_reads <- OTU_reads_total(OTU_V1)
V6_total_reads <- OTU_reads_total(OTU_V6)
ITS1_total_reads <- OTU_reads_total(OTU_ITS1)
total_reads <- V4_total_reads + V1_total_reads + V6_total_reads + ITS1_total_reads

# read rarefied data
rarefy_V4 <- read.csv(here::here("analysis","data","derived_data", "V4_shared_format.1.subsample.60216.csv"), sep = ",")
rarefy_V1 <- read.csv(here::here("analysis","data","derived_data","V1_shared_format.1.subsample.13658.csv"), sep = "")
rarefy_V6 <- read.csv(here::here("analysis","data","derived_data", "V6_shared_format.1.subsample.28402.csv"), sep = ",")
rarefy_ITS1 <- read.csv(here::here("analysis","data","derived_data", "ITS1_shared_format.1.subsample.387.csv"), sep = ",")

rarefy_reads <- function(df) {
  sum_read <- df %>%
    rowwise() %>% 
    mutate(sum_read = sum(across(starts_with("ASV")))) %>% 
    select(sum_read)
  return(sum(sum_read))
}

V4_rare_reads <- rarefy_reads(rarefy_V4)
V1_rare_reads <- rarefy_reads(rarefy_V1)
V6_rare_reads <- rarefy_reads(rarefy_V6)
ITS1_rare_reads <- rarefy_reads(rarefy_ITS1)
rare_reads <- V4_rare_reads + V1_rare_reads + V6_rare_reads + ITS1_rare_reads

V4_OTU_num <- unique(rarefy_V4$numOtus)
V1_OTU_num <- unique(rarefy_V1$numOtus)
V6_OTU_num <- unique(rarefy_V6$numOtus)
ITS1_OTU_num <- unique(rarefy_ITS1$numOtus)
total_OTU <- V4_OTU_num + V1_OTU_num + V6_OTU_num + ITS1_OTU_num
```

## Alpha diversity analysis of bacterial and fungal communities
Illumina analysis generated a total of `r total_reads` sequences that were normalized to `r rare_reads` after rarefaction [@Schloss2024]. Soil samples have the highest shannon diversity index and richness, followed by pottery exterior samples, and pottery interior samples, and negative control.

## Taxonomic composition

## Beta diversity analysis

Alpha diversity 
Beta diversity
Relative abundance 
Venn diagrams
The abundant taxa at ASV level & the abundant taxa at genus level

Questions:

1. Abundance and relative abundance? Proportion of the general microbial communities across different types of analysis. Richness and Shanon analysis
2. What should I focus on heatmap? those high frequencies? targets? or based on tree? 
3. How to deal with control sets? Subtract? 
4. What is the correction between V1, V4, V6 of 16S according to abundance and taxonomy?
5. Phylogentic-based analysis, phylogenic tree?
6. the distribution of microbiome across different groups of samples?

ANOVA
normal distribution
standard deviation 
Kristall_willis

<!-- Here's some example analysis code: -->

```{r}
#| label: fig-demo-plot
#| fig-cap: "A plot of random numbers"
plot(rnorm(10))
```

@fig-demo-plot shows how we can have a caption and cross-reference for a plot. Note that figure label and cross-references must both be prefixed with `fig-`

```{r}
#| label: demo-inline-code
#| echo: false
x <- round(pi, 2)
```

Here is an example of inline code `r x` in the middle of a sentence.

# Discussion

There are significant differences in the richness and diversity of bacterial and archaeal communities across the four groups: soil > exterior potsherds > interior potsherds > control
Some microbial groups are significantly more abundant in pottery samples, such as Weissella, Escherichia Shigella, Paenibacillus, and Chryseobacterium, and Penicillium
V4 and V6 regions are more ideal for determining the microorganisms in archaeological materials compared with V1 and ITS1 regions

# Conclusion

# Acknowledgements

<!-- The following line inserts a page break  -->

\newpage

# References

<!-- The following line ensures the references appear here for the MS Word or HTML output files, rather than right at the end of the document (this will not work for PDF files):  -->

::: {#refs}
:::

\newpage

### Colophon

This report was generated on `r Sys.time()` using the following computational environment and dependencies:

```{r}
#| label: colophon
#| cache: false

# which R packages and versions?
if ("devtools" %in% installed.packages()) devtools::session_info()
```

The current Git commit details are:

```{r}
# what commit is this file at? 
if ("git2r" %in% installed.packages() & git2r::in_repository(path = ".")) git2r::repository(here::here())  
```