### 2024-02-05 analysis/supplementary-materials/Supplement.qmd
work on supplement

---
title: "Supplementary Materials for Characterization of the microbial and fungal communities in ancient pottery from Neolithic southwestern China"
author:
  - Li-Ying Wang:
      correspondence: "yes"
      email: liying15@gmail.com
      orcid: 0000-0001-9060-2125
      institute:
        - AS
  - Kuei-Chen Lin:
      institute: AS
      orcid: 0000-0003-1689-0558
  - Zhiqing Zhou:
      orcid: 0000-0003-1689-0551
      institute:
        - name: Chengdu Institute of Cultural Relics and Archaeology
  - Yin-Ru Chiang:
      correspondence: "yes"
      institute: AS
      orcid: 0000-0003-1689-0559
institute:
  - AS:
      name: Academia sinica
      address: 130, Section 2, Academia Rd, Nangang District, Taipei City, Taiwan
title-block-published: "Last updated"  
date: now
date-format: long
format: 
  docx:
    reference-doc: "../templates/template.docx" # Insert path for the DOCX file
execute:
  echo: true
  warning: false
  message: false
  comment: "#>"
  fig-path: "../figures/"
  fig-dpi: 600
filters:
  - ../templates/scholarly-metadata.lua
  - ../templates/author-info-blocks.lua
  - ../templates/pagebreak.lua
csl: "../templates/journal-of-archaeological-science.csl" # Insert path for the bib-style
---

## Alpha Diversity

```{r get-data}
#| label: get-data
#| echo: false
#| cache: true
library(tidyverse)
library(readxl)
# read rarefied data
rarefy_V4 <- read.csv(here::here("analysis","data","derived_data", "V4_shared_format.1.subsample.60216.csv"), sep = ",")
rarefy_V1 <- read.csv(here::here("analysis","data","derived_data","V1_shared_format.1.subsample.13658.csv"), sep = "")
rarefy_V6 <- read.csv(here::here("analysis","data","derived_data", "V6_shared_format.1.subsample.28402.csv"), sep = ",")
rarefy_ITS1 <- read.csv(here::here("analysis","data","derived_data", "ITS1_shared_format.1.subsample.387.csv"), sep = ",")
context <- read_excel(here::here("analysis","data", "raw_data", "GJB_amp_meta.xlsx"))
```

```{r}
#| label: alpha-richness
#| echo: false
#| cache: true
library(knitr)
library(vegan) 
# to get the same order for all
rarefy_V4 <- rarefy_V4 %>% mutate(Group = tolower(Group)) %>% arrange(Group)
rarefy_V1 <- rarefy_V1 %>% mutate(Group = tolower(Group)) %>% arrange(Group)
rarefy_V6 <- rarefy_V6 %>% mutate(Group = tolower(Group)) %>% arrange(Group)
rarefy_ITS1 <- rarefy_ITS1 %>% mutate(Group = tolower(Group)) %>% arrange(Group)

# tidy meta data and get the same order as the rarefy data
context_tidy <-
  context %>%
  mutate(`adpater no` = tolower(`adpater no`),
         type = tolower(str_extract(`adpater no`, "[[:alpha:]]+(?=\\d)")),
         type = case_when(type == "con" ~ "control", TRUE ~ as.character(type)),
         area = case_when(area == "inside" ~ "interior", area == "outside" ~ "exterior"),
         group = if_else(!is.na(area)&!area%in% c("belly", "outside belly"),
                         paste0(type, "-", area), type),
         group = factor(group, levels = c("control","pot-interior","pot-exterior","soil"))) %>%
  arrange(`adpater no`)

# function to have OTU reads only
reads_only <- function(df) {
  df %>%
    select(-label, -Group, -numOtus)
}

rarefy_V4_reads_only  <- reads_only(rarefy_V4)
rarefy_V1_reads_only  <- reads_only(rarefy_V1)
rarefy_V6_reads_only  <- reads_only(rarefy_V6)
rarefy_ITS1_reads_only  <- reads_only(rarefy_ITS1)

# function to get context for each region
meta_region <- function(df, region){
  context_tidy %>%
    filter(Region == region) %>%
    arrange(`adpater no`)
}

V4_meta<- meta_region(context_tidy, "V4")
V1_meta<- meta_region(context_tidy, "V1")
V6_meta<- meta_region(context_tidy, "V6")
ITS1_meta<- meta_region(context_tidy, "ITS1")

# functions calculate richness, Chao1, evenness index, and Shannon index 
alpha_cal_df <- function(df1, df2){
  richness <- estimateR(df1)
  evenness <- diversity(df1)/log(specnumber(df1))
  shannon <- diversity(df1, index = "shannon")
  alphadiv <- cbind(df2, t(richness), shannon, evenness) #combine all in one
  return(alphadiv)
}

V4_alphadiv <- alpha_cal_df(rarefy_V4_reads_only, V4_meta)
V1_alphadiv <- alpha_cal_df(rarefy_V1_reads_only, V1_meta)
V6_alphadiv <- alpha_cal_df(rarefy_V6_reads_only, V6_meta)
ITS1_alphadiv <- alpha_cal_df(rarefy_ITS1_reads_only, ITS1_meta)

# pairwise Wilcoxon rank-sum  on richness and shannon
library(rstatix)

rich_pair_V4 <- V4_alphadiv %>% pairwise_wilcox_test(S.obs ~ group, p.adjust.method = 'fdr')
rich_pair_V1 <- V1_alphadiv %>% pairwise_wilcox_test(S.obs ~ group, p.adjust.method = 'fdr')
rich_pair_V6 <- V6_alphadiv %>% pairwise_wilcox_test(S.obs ~ group, p.adjust.method = 'fdr')
rich_pair_ITS1 <- ITS1_alphadiv %>% pairwise_wilcox_test(S.obs ~ group, p.adjust.method = 'fdr')

# function to add region name
rich_pair_region <- function(df, vector) {
  df %>%
    mutate(region = vector) %>% 
    select(region, group1, group2, n1, n2, p, p.adj, p.adj.signif)
}

rich_pair_V4 <- rich_pair_region(rich_pair_V4, "V4")
rich_pair_V1 <- rich_pair_region(rich_pair_V1, "V1")
rich_pair_V6 <- rich_pair_region(rich_pair_V6, "V6")
rich_pair_ITS1 <- rich_pair_region(rich_pair_ITS1, "ITS1")

rich_pair_all <-
  rbind(rich_pair_V1, rich_pair_V4, rich_pair_V6, rich_pair_ITS1)

kable(rich_pair_all , caption = "Table S1: Pairwise comparison of Richness index (the number of observed OTUs) using Wilcoxon rank-sum test with FDR-adjusted p-value")
```

```{r}
#| label: alpha-shannon
#| echo: false
#| cache: true

# pairwise t-test on Shannon
shan_pair_V4 <- V4_alphadiv %>% pairwise_wilcox_test(shannon ~ group, p.adjust.method = 'fdr')
shan_pair_V1 <- V1_alphadiv %>% pairwise_wilcox_test(shannon ~ group, p.adjust.method = 'fdr')
shan_pair_V6 <- V6_alphadiv %>% pairwise_wilcox_test(shannon ~ group, p.adjust.method = 'fdr')
shan_pair_ITS1 <- ITS1_alphadiv %>% pairwise_wilcox_test(shannon ~ group, p.adjust.method = 'fdr')

# function to add region name
shan_pair_region <- function(df, vector) {
  df %>%
    mutate(region = vector) %>% 
    select(region, group1, group2, n1, n2, p, p.adj, p.adj.signif)
}

shan_pair_V4 <- shan_pair_region(shan_pair_V4, "V4")
shan_pair_V1 <- shan_pair_region(shan_pair_V1, "V1")
shan_pair_V6 <- shan_pair_region(shan_pair_V6, "V6")
shan_pair_ITS1 <- shan_pair_region(shan_pair_ITS1, "ITS1")

shan_pair_all <-
  rbind(shan_pair_V1, shan_pair_V4, shan_pair_V6, shan_pair_ITS1)

kable(shan_pair_all  , caption = "Table S2: Pairwise comparison of Shannon diversity indices using Wilcoxon rank-sum test with FDR-adjusted p-value")
```

## Including Plots

You can also embed plots, for example:

```{r pressure, echo=FALSE}
plot(pressure)
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.